\documentclass{beamer}

\newcommand{\week}{Financial Markets 1 of 8}

\title{Financial Market Returns}
\subtitle{Mishkin Chapter 4}
\author{Econ 357}
\date{\week}

% Reference the shared preamble
\input{preamble}

% Set the footer -- change 
\setbeamertemplate{footline}{
  \leavevmode%
  \vspace{2ex}
  \hbox{%
    % Left box: Econ 457
    \begin{beamercolorbox}[wd=.4\paperwidth,ht=2.5ex,dp=1ex,left]{author in head/foot}%
      \hspace{1em}Econ 357
    \end{beamercolorbox}%
    % Middle box: Week
    \begin{beamercolorbox}[wd=.2\paperwidth,ht=2.5ex,dp=1ex,center]{date in head/foot}%
      \centering\week
    \end{beamercolorbox}%
    % Right box: Slide numbers
    \begin{beamercolorbox}[wd=.4\paperwidth,ht=2.5ex,dp=1ex,center]{date in head/foot}%
      \hfill\insertframenumber{} 
    \end{beamercolorbox}%
  }%
  \vskip0pt%
}

\begin{document}

\frame{\titlepage}

\begin{frame}[t]
    \frametitle{Review}
        \textbf{Intro 1}
            \begin{itemize}
                \item Real Assets and Financial assets.
                \item Every financial assets is somebody's asset and somebody else's liability.
            \end{itemize}

        \vspace{1em}

        \textbf{Intro 2}
            \begin{itemize}
                \item Financial Intermediaries
                \item Reasons they exist: transaction costs, risk sharing, Asymmetric Information.
                \item Asymmetric Information: Adverse Selection and Moral Hazard.
            \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{Outline for Financial Markets}
        \begin{enumerate}
            \item \fbox{Financial Market Returns}
            \item Bonds 1: Discounting, Prices and Yields
            \item Bonds 2: Nominal v. Real Yields, Supply and Demand
            \item Bonds 3: Market for Money
            \item Bonds 4: Risk and Term Structure
            \item Equities 1: Dividend Discount Model
            \item Equities 2: P/E Ratio
            \item Equities 3: Theories of Stock Pricing
            
        \end{enumerate}
\end{frame}


\begin{frame}
    \frametitle{Outline for Today's Lecture}
    \begin{enumerate}
        \item Financial Market Returns
        \begin{itemize}
            \item Total Return Definition
            \item Measurement
        \end{itemize}
        \item Real Returns
        \item History
        \item Reading - Black Swans
    \end{enumerate}
\end{frame}

\begin{frame}[t]
    \frametitle{1. Financial Market Returns}
    \framesubtitle{Total Returns}

    Definition of \textbf{total return}:
    $$\text{Return (\%)} = \frac{\text{Income}}{\text{Initial Price}} + \frac{\text{Change in Price}}{\text{Initial Price}} $$

    \vspace{1em}
    Notes:
    \begin{itemize}
        \item This is a \textit{backwards looking} formula.   
        \item Calculating total return requires three values: beginning and ending price and income.   Given total return and two of the three values, could calculate the third value by rearranging the equation.
        \item Expressed as a percentage of the initial price.
        \item Don't forget income!  Only include income that was earned during the holding period.
    \end{itemize}  
\end{frame}

\end{document}